---
title: Car Insurance Claims Classification Report
authors:
    name: Hamed Vaheb
    email: hamed.vaheb.001@student.uni.lu
abstract: 
  This report describes the machine learning project which aims at analysis and prediction of the car insurance claims dataset. Based on historical data of the clients, the variable of interest is the outcome of insurance, which indicates whether a customer has claimed his loan or not. Two machine learning models, i.e., logistic regression and random forest is implemented to predict the outcomes using the historical data. The results indicate successful and promising predictions with accuracy of over 80 percent. Besides, prediction, informative insights from the dataset and models have been drawn. For instance, by dint of extracting feature importance from the random forest model, the parts of the dataset that play the more effective role in the prediction become evidenced.

bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Introduction
The use of statistical learning models has been a common practice in actuarial science since the 1980s. 
It was not long after since the field adoped linear models and generalized linear models (GLMs).
However, it was in the past few years that the use of AI and machine learning garnered significant attention in the field \cite{rev1}.


# Headings: first level
\label{sec:headings}


LaTeX command can be used to reference other section. See Section \ref{sec:headings}.
However, you can also use **bookdown** extensions mechanism for this.

## Headings: second level

You can use equation in blocks

$$
\xi _{ij}(t)=P(x_{t}=i,x_{t+1}=j|y,v,w;\theta)= {\frac {\alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}{\sum _{i=1}^{N} \sum _{j=1}^{N} \alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}}
$$

But also inline i.e $z=x+y$

### Headings: third level

Another paragraph. 


# Examples of citations, figures, tables, references
\label{sec:others}

You can insert references. Here is some text [@kour2014real; @kour2014fast] and see @hadash2018estimate.

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}


## Figures

You can insert figure using LaTeX directly. 

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.


## Tables

Below we can see how to use tables. 

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```


## Lists

- Item 1
- Item 2 
- Item 3
